---
title: "CDF_endophytes"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
author: "Briana K. Whitaker"
date: "`r Sys.Date()`"
---
\fontsize{9}{10}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning =FALSE, message=FALSE)
```
---


* Run using `r version[['version.string']] `.



```{r, echo=FALSE, results='hide', include=FALSE}
#setwd("C:/Users/Briana.Whitaker/OneDrive - USDA/2-ResearchProjects/2022-CornDebrisFungicides")

x<-c("dplyr",  "ggplot2", "vegan" , "car", "lme4", "viridis", "RRPP",
     "grDevices", "tidyr", "indicspecies")
lapply(x, require, character.only = TRUE)
#require(nlme)

#Load functions
source("./multiplot.function.R")
`%nin%` <- Negate(`%in%`)

set.seed(407)

# function for ellipsess - 
veganCovEllipse <- function (cov, center = c(0, 0), scale = 1, npoints = 100)   {
  theta <- (0:npoints) * 2 * pi/npoints
  Circle <- cbind(cos(theta), sin(theta))
  t(center + scale * t(Circle %*% chol(cov)))    }

myMEAN <- function(x) {
    x1 <- na.omit(x)  #number of omitted values length(attr(test,"na.action"))
    MEAN <- mean(x1)
}
mySD <- function(x) {
    x1 <- na.omit(x)  #number of omitted values length(attr(test,"na.action"))
    SD <- sd(x1)
}

myREP <- function(x) {
    x1 <- na.omit(x)  #number of omitted values length(attr(test,"na.action"))
    REP <- length(x1) #number of values used in mean
}


# set ggplot2 theme
theme_set(theme_bw(base_size=14)) 
theme_update(panel.grid.major=element_line(0), panel.grid.minor=element_line(0))


trt_color <- c("#66C2A5","#66C2A5", "#FC8D62", "#FC8D62")
fung_color <- c("#8DA0CB", "#E78AC3")
```



# Load Data
```{r}
plot <- read.csv("./Data/CornDebrisFungicides_ExperimentalDesign-Plots.csv",
                 stringsAsFactors = TRUE)
samps <- read.csv("./Data/CornDebrisFungicides_ExperimentalDesign-CultureSamples.csv",
                 stringsAsFactors = TRUE)   
                 #*NOTE -mismatch in orig names C208 vs. C-208, manually fixed
isols <- read.csv("./Data/CornDebrisFungicides_ExperimentalDesign-Isolates-2023-06-27.csv",
                 stringsAsFactors = TRUE)
                 #*NOTE - problem with naming of sample_ID in this sheet C208 
                 #*rather than C-208, etc. had to manually rename
taxon <- read.csv("./Data/Sequencing/ab1/ITSx/CDF_OTU97_cons_RDP_Combined.csv",
                  row.names = 1)   

```

```{r, include = FALSE}
# add OTU info to taxon file
#key <- isols %>% select(OTU, contig.name) %>% unique
#taxon$OTU <- key$OTU[match(taxon$Contig, key$contig.name)]


#merge trt info
samps <- merge(samps, plot, by=c("PlotID"), all.x=TRUE)
isols <- merge(isols, samps, by="Sample_ID", all.x = TRUE)

# add sample level IF
IF <- as.data.frame(table(isols$Sample_ID))
identical(sort(IF$Var1), sort(samps$Sample_ID))
samps <- merge(samps, IF, by.x = "Sample_ID", by.y = "Var1", all.x = TRUE)
samps$IF <- samps$Freq/32

# add treatment combo
samps$Trt <- as.factor(paste(samps$Site, samps$FungicideTrt, sep = "-"))
levels(samps$Trt)[1] <- "Champaign-Control"
levels(samps$Trt)[2] <- "Champaign-Delaro"
levels(samps$Trt)[3] <- "Gilman-Control"
levels(samps$Trt)[4] <- "Gilman-Delaro"

# make factor
samps$PlotNum <- as.factor(samps$PlotNum)

# Create SbyS
table(isols$sequenced, isols$Site)
(252+25+11+0)/(294)  #98.0% of the data for Champaign site
(84+36+13+2)/(137)     #98.5% of the data for Gilman site
# 6 failed and 2 failed to amplify from Champ and Gilman

#subset to completed seqs
done <- isols %>% filter(isols$sequenced == "yes-pt1" | 
                         isols$sequenced == "yes-pt2" |
                         isols$sequenced == "yes-pt3" |
                         isols$sequenced == "yes-pt4" )

done <- droplevels(done)
```


### Make SbyS and Calculate Diversity
```{r, include = FALSE}
# create SbyS (completed sequences)
SbyS <- as.matrix(table(done$Sample_ID , done$OTU))

# rename column headers by taxon name
colnames(SbyS) <- paste(colnames(SbyS), 
    taxon$GenBankName[match(colnames(SbyS), taxon$OTU)],
    sep = ":")

# without Singletons
singles <- colSums(SbyS)
to.remove <- names(singles[singles==1 | singles == 0])
SbyS.noSing <- SbyS[,colnames(SbyS) %nin% to.remove] 


# Hellinger transformation
SbyS.hel <- as.data.frame.matrix(decostand(SbyS.noSing, method = "hellinger")) 
            # samples are rows (MARGIN = 1 default)
rownames(samps) <- samps$Sample_ID
identical((rownames(SbyS.hel)), (rownames(samps))) # sanity check
rowSums(SbyS.hel)




# compute diversity info
Sdat <- data.frame(specnumber(SbyS), diversity(SbyS, index = "shannon"),
                   diversity(SbyS, index = "invsimpson"))
rownames(Sdat) -> Sdat$Sample_ID
colnames(Sdat) <- c("Sdiv", "Hdiv", "Simp", "Sample_ID")
#plot(Hdiv ~ Sdiv, data = Sdat)
#plot(Hdiv ~ Simp, data = Sdat)

# merge data
samps <- merge(samps, Sdat, by="Sample_ID", all.x = TRUE)
samps$Site_short <- sapply(samps$Site, function(x) substring(x, 1, 1))
rownames(samps) <- samps$Sample_ID


# create a means dataset for making barplots
samp_means <- data.frame(
    Trt=c(levels(samps$Trt)),
    Site = c(rep(levels(samps$Site), each = 2)),
    FungicideTrt=c(rep(levels(samps$FungicideTrt),2)),

    #IF
    IF_mean=as.vector(tapply(samps$IF, list(samps$Trt), myMEAN)),
    IF_sd=as.vector(tapply(samps$IF, list(samps$Trt), mySD)),
    IF_rep=as.vector(tapply(samps$IF, list(samps$Trt), myREP)),
    
    #Sdiv
    Sdiv_mean=as.vector(tapply(samps$Sdiv, list(samps$Trt), myMEAN)),
    Sdiv_sd=as.vector(tapply(samps$Sdiv, list(samps$Trt), mySD)),
    Sdiv_rep=as.vector(tapply(samps$Sdiv, list(samps$Trt), myREP)),
    
    #Hdiv
    Hdiv_mean=as.vector(tapply(samps$Hdiv, list(samps$Trt), myMEAN)),
    Hdiv_sd=as.vector(tapply(samps$Hdiv, list(samps$Trt), mySD)),
    Hdiv_rep=as.vector(tapply(samps$Hdiv, list(samps$Trt), myREP))
    )

### compute singleton info
SbyS.onlySing <- SbyS[,colnames(SbyS) %in% to.remove]
samps$Singles <- rowSums(SbyS.onlySing)


```

### Experimental Design
```{r}
table(samps$Site, samps$FungicideTrt)
table(samps$Site, samps$PlotID) #nested
table(samps$PlotID, samps$FungicideTrt, samps$Site) #nested

```



# Descriptive stats for MS
```{r}
dim(SbyS.noSing)
dim(SbyS)
# 14 non singleton OTUs, 32 OTUs total (so 18 singleton OTUs)


samps %>% 
  group_by(Site) %>%  #, FungicideTrt
  summarize(mean = myMEAN(IF), sd = mySD(IF)) %>% 
  ungroup %>% as.data.frame -> IF_summ
IF_summ

table(isols$Site)
table(isols$FungicideTrt)

#table(isols$OTU, isols$Site)>0
#table(isols$OTU, isols$FungicideTrt)>0

table(isols$contig.name)

```

# Other Figures
### Rarefaction Curves
```{r, echo = FALSE, fig.width = 4}
# Make subsets
champ.ids <- samps$Sample_ID[samps$Site=="Champ-Kleczewski"]
gilman.ids <- samps$Sample_ID[samps$Site=="Gilman-Kleczewski"]
champ.site <- SbyS[(rownames(SbyS) %in% champ.ids),]
gilman.site <- SbyS[(rownames(SbyS) %in% gilman.ids),]

cont.ids <- samps$Sample_ID[samps$FungicideTrt =="Control"]
fung.ids <- samps$Sample_ID[samps$FungicideTrt=="Delaro Complete 12oz"]
cont.trt <- SbyS[(rownames(SbyS) %in% cont.ids),]
fung.trt <- SbyS[(rownames(SbyS) %in% fung.ids),]

#make curves
x.full <- specaccum(SbyS, method='random', permutations=999, gamma="bootstrap")
x.collector <- specaccum(SbyS, method='collector', permutations=999, gamma="bootstrap")

x.champ <- specaccum(champ.site, method='random', permutations=999, gamma="bootstrap")
x.gilman <- specaccum(gilman.site, method='random', permutations=999, gamma="bootstrap")

x.cont <- specaccum(cont.trt, method='random', permutations=999, gamma="bootstrap")
x.fung <- specaccum(fung.trt, method='random', permutations=999, gamma="bootstrap")

# make data formatted for abundance plot
total <- sum(colSums(SbyS))
otu_Count <- data.frame(sort(colSums(SbyS), dec = TRUE))
colnames(otu_Count) <- "Count"
otu_Count$pi <- otu_Count$Count/total
otu_Count$otu <- rownames(otu_Count)
otu_Count$otu <- factor(otu_Count$otu, levels = rownames(otu_Count))
otu_Count <- otu_Count %>% filter(Count > 3)

# # Panel A - Abundance Plot
# abund_plot <- ggplot(otu_Count, aes(x = otu, y = pi)) +
#     geom_bar(stat = 'identity', fill = "navyblue", colour = 'black') +
#     theme(axis.text.x = element_text(angle = 30, hjust = 0.95)) +
#     scale_y_continuous("Relative\nAbundance", expand = c(0, 0)) +
#     scale_x_discrete("") +
#     labs(subtitle='(a)') +
#     theme(axis.text = element_text(size = 10), axis.title.y = element_text(size = 11),
#           axis.title.x = element_blank(), plot.subtitle=element_text(size=11, face = 'bold')) 
# abund_plot


#tiff("./Fig2_rarecurves.tiff", width=3.25, height=8, units="in", res=600)
#change native R graphics to Helvetica font
windowsFonts(A = windowsFont("Helvetica")) #, "Helvetica Bold", "Helvetica Oblique","Helvetica Bold Oblique"
# Potential panel figure
par(mfrow = c(3,1), mar=c(4.1,4.1,1.1,0.4), oma=c(0,0,0,0), family = "A")  #bot,L,top,R
# # panel A - leave empty
# plot.new() 
# # panel A - native R plot
plot(x.full, add=FALSE, lwd=2, ci.type="line", ci.col="black", ci.lty=2, 
     xlab="No. of Plant Samples", ylab="Accumulated Taxa", xlim=c(0,36))
plot(x.collector, add=TRUE, lwd=2, col="red")
legend(24,18, bty="n", c("Estimated S","Observed S","95% CIs"),
       lty=c(1,1,2),col=c("black","red","black"),lwd=1.3, cex=0.9)
title(main="(a)", adj=0, cex.main=1.4)

# panel B - native R plot
plot(x.cont, add=FALSE, lwd=3, ci=0, lty=1, col="#1B9E77",
     xlab="No. of Plant Samples", ylab="Accumulated Taxa", xlim=c(0,18), ylim=c(0,25))
plot(x.gilman, add=TRUE, lwd=3, ci=0, lty=2, col="#D95F02")
legend(12,9.5, bty="n", c("Champaign","Gilman"),
       lty=c(1:3),col=c("#1B9E77","#D95F02"), lwd=1.3, cex=0.9)
title(main="(b)", adj=0, cex.main=1.4)

# panel C - native R plot
plot(x.champ, add=FALSE, lwd=3, ci=0, lty=1, col="#8DA0CB",
     xlab="No. of Plant Samples", ylab="Accumulated Taxa", xlim=c(0,18), ylim=c(0,25))
plot(x.fung, add=TRUE, lwd=3, ci=0, lty=2, col="#E78AC3")
legend(12.5,9.5, bty="n", c("Control","Delaro"),
       lty=c(1:3),col=c("#8DA0CB","#E78AC3"), lwd=1.3, cex=0.9)
title(main="(c)", adj=0, cex.main=1.4)

# # overlay panel A - using grid package, 
# vp <- grid::viewport(height = unit(1/4,"npc"), width=unit(1, "npc"), 
#               just = "top",
#               y = 1, x = 0.5)
# print(abund_plot, vp = vp)

#dev.off()
```


### Rel. Abund. Plot
```{r, echo = FALSE, results = 'hide'}
top <- sort(colSums(SbyS), decreasing = TRUE)[1:9]
topASVs <- names(top)
otherASVs <- c(colnames(SbyS)[colnames(SbyS) %nin% topASVs])

#identical(rownames(samps), rownames(SbyS)) #must be TRUE

#merge sample data with rel.abundance of OTUs
out <- samps  %>% select(Trt, Site, FungicideTrt)
SbyStot <- decostand(SbyS,"total")[,topASVs]
Other <- rowSums(decostand(SbyS,"total")[,otherASVs])
SbyS_rel <- cbind(SbyStot, Other)
dat <- cbind(out, SbyS_rel)
# covert to long-format
dat2 <- reshape2::melt(dat, id.vars=c("Trt", "Site", "FungicideTrt"))

# make an ASV color variable, with gray for 'other'
# asv_color <- c(viridis_pal(option = "plasma")(length(levels(dat2$variable))-1),
#                "#696969")
asv_color <- c("#e60049", "#0bb4ff", "#50e991", "#e6d800", "#9b19f5", 
               "#ffa300", "#dc0ab4", "#b3d4ff", "#00bfa0",     "#696969") 

trt.dat = dat2 %>% group_by(Trt, variable) %>% summarize(value = sum(value))
a <- ggplot(trt.dat, aes(x = Trt, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = 'fill') +
  scale_fill_manual("Fungi Names", values = asv_color) +
  xlab("") + ylab("Relative Abundance of Fungal Endophytes") +
  #guides(fill = guide_legend(nrow=5, byrow=TRUE)) +
  theme_bw() +
  theme(axis.text.x  = element_text(angle=55, hjust = .95,  size=10),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 9),
        legend.margin = margin(c(0,5,0,0)),
        plot.margin = unit(c(1,1,1,1), "mm")) 
  #legend.position = 'bottom'
  #theme(axis.title.x = element_text(size=8),
  #      axis.text.x  = element_text(size=8)) +
#tiff("./Fig1_RelAbun_trt.tiff", width=3.25, height=5, units="in", res=600)
a
#dev.off()
```


### OTU Violin plots
```{r, include = FALSE}
otu_plot <- ggplot(dat2[dat2$variable!="Other",], aes(x = Trt, y = value, fill = Site)) +
    facet_grid(variable~., scales = "free") +
    geom_violin() +
    geom_jitter(width=.1, height = 0) +
    scale_fill_manual("Names", values = trt_color[c(1,3)]) +
    scale_x_discrete("Location & Fungicide Treatment") + 
    scale_y_continuous("Relative Abundance of Different Fungi per Plant") +
    theme(axis.text.x  = element_text(angle=60, vjust=0.5, size=10)) +
    guides(fill = "none")
#tiff("./OTUs_trt.tiff", width=4, height=10, units="in", res=600)
otu_plot
#dev.off()
```

# Alpha Diversity & IF
### IF
```{r}
par(mfrow = c(1,1))
hist(samps$IF)
par(mfrow = c(2,2))

# Isolation Frequency
if1 <- lm(IF ~ Site_short*FungicideTrt, data=samps,      
          #should really have plot as random int. ?
                 na.action=na.omit)
Anova(if1, type = 3)
summary(if1)
plot(if1)

# if1a <- lmer(IF ~ Site_short*FungicideTrt + (1|PlotID), data=samps,      
#                  na.action=na.omit)
# # isSingular model, not able to estimate PlotID residuals
# Anova(if1a, type = 3)
# summary(if1a)  #PlotID variance is 0, that's why singular model



# if1b <- lm.rrpp(IF ~  Site_short + FungicideTrt +
#                       Site_short:FungicideTrt + Site_short:FungicideTrt:PlotID,
#                  data = samps, iter = 999, print.progress = FALSE, SS.type="III")   
# #anova(rrpp1a, effect.type = "F")
# anova(if1b, effect.type = "F", 
#      error = c("Site_short:FungicideTrt:PlotID", "Site_short:FungicideTrt:PlotID",
#                "Site_short:FungicideTrt:PlotID", "Residuals"))
#                                Df      SS       MS     Rsq      F        Z Pr(>F)  
# Site_short                      1 0.08610 0.086100 0.06294 7.1007  2.01014  0.018 *
# FungicideTrt                    1 0.04704 0.047038 0.03439 3.8792  1.60837  0.043 *
# Site_short:FungicideTrt         1 0.01172 0.011719 0.00857 0.9664  0.53871  0.311  
# Site_short:FungicideTrt:PlotID  8 0.09701 0.012126 0.07091 0.4817 -1.13353  0.873  
# Residuals                      24 0.60417 0.025174 0.44167                         
#summary(if1b)


range(samps$IF)
tapply(samps$IF, samps$Site, mean)

```
### Richness
```{r}
par(mfrow = c(2,2))

# Richness
rich1 <- glm(Sdiv ~ Site_short*FungicideTrt, data=samps,      
             #should really have plot as random int. ?
                 family=poisson, na.action=na.omit)
Anova(rich1, type = 3)
summary(rich1)
plot(rich1)
#rich1$deviance/rich1$df.residual #0.63
# dispersionStat <- function (x) 
# {
#     pr <- sum(residuals(x, type = "pearson")^2)
#     dispersion <- pr/x$df.residual
#     return(c(pearson.chi2 = pr, dispersion = dispersion))
# }
# dispersionStat(rich1) # dispersion 0.6215236


# rich2 <- glmer(Sdiv ~ Site_short*FungicideTrt + (1|PlotID), data=samps,      
#                  family=poisson, na.action=na.omit)
# # result isSIngular = TRUE
# Anova(rich2, type = 3)  
# #summary(rich2)   #can't estimate plot intercept effect, variance = 0

# rich2a <- lm.rrpp(Sdiv ~  Site_short + FungicideTrt +
#                       Site_short:FungicideTrt + Site_short:FungicideTrt:PlotID,
#                  data = samps, iter = 999, print.progress = FALSE, SS.type="III")   
# #anova(rrpp1a, effect.type = "F")
# anova(rich2a, effect.type = "F", 
#      error = c("Site_short:FungicideTrt:PlotID", "Site_short:FungicideTrt:PlotID",
#                "Site_short:FungicideTrt:PlotID", "Residuals"))
#                                Df      SS     MS     Rsq      F        Z Pr(>F)
# Site_short                      1   6.000 6.0000 0.05192 2.5116  1.22566  0.116
# FungicideTrt                    1   2.667 2.6667 0.02308 1.1163  0.69218  0.255
# Site_short:FungicideTrt         1   0.083 0.0833 0.00072 0.0349 -1.06026  0.841
# Site_short:FungicideTrt:PlotID  8  19.111 2.3889 0.16538 0.9053 -0.03022  0.520
# Residuals                      24  63.333 2.6389 0.54808                       
# Total                          35 115.556                                      


range(samps$Sdiv)
tapply(samps$Sdiv, samps$Site, mean)

```


### Shannon Diversity
```{r}
par(mfrow = c(1,1))
hist(samps$Hdiv)
par(mfrow = c(2,2))


# Diversity
div1 <- lm(Hdiv ~ Site_short*FungicideTrt, data=samps, na.action=na.omit)
Anova(div1, type = 3)
summary(div1)
plot(div1)


# div2 <- lmer(Hdiv ~ Site_short*FungicideTrt + (1|PlotID), data=samps, 
#              na.action=na.omit)
# # result isSIngular = TRUE
# Anova(div2, type = 3)  
# #summary(div2)   #can't estimate plot intercept effect, variance = 0

# div2a <- lm.rrpp(Hdiv ~  Site_short + FungicideTrt +
#                       Site_short:FungicideTrt + Site_short:FungicideTrt:PlotID,
#                  data = samps, iter = 999, print.progress = FALSE, SS.type="III")   
# #anova(rrpp1a, effect.type = "F")
# anova(div2a, effect.type = "F", 
#      error = c("Site_short:FungicideTrt:PlotID", "Site_short:FungicideTrt:PlotID",
#                "Site_short:FungicideTrt:PlotID", "Residuals"))
#                                Df     SS       MS     Rsq      F        Z Pr(>F)
# Site_short                      1 0.2032 0.203191 0.03156 1.4346  0.83909  0.212
# FungicideTrt                    1 0.1759 0.175895 0.02732 1.2419  0.75485  0.234
# Site_short:FungicideTrt         1 0.0301 0.030142 0.00468 0.2128 -0.35328  0.640
# Site_short:FungicideTrt:PlotID  8 1.1331 0.141636 0.17599 0.9251  0.01883  0.496
# Residuals                      24 3.6746 0.153109 0.57073                       
# Total                          35 6.4385                                        


range(samps$Hdiv)
tapply(samps$Hdiv, samps$Site, mean)
tapply(samps$Hdiv, samps$FungicideTrt, mean)
```

### Inverse Simpson
```{r}
par(mfrow = c(1,1))
hist(sqrt(samps$Simp))
par(mfrow = c(2,2))

# Evenness
simp1 <- lm(log(Simp) ~ Site_short*FungicideTrt, data=samps, na.action=na.omit)
Anova(simp1, type = 3)
summary(simp1)
plot(simp1)  #not great


# simp2 <- lmer(sqrt(Simp) ~ Site_short*FungicideTrt + (1|PlotID), data=samps, 
#              na.action=na.omit)
# # result isSIngular = TRUE
# Anova(simp2, type = 3)  
# #summary(simp2)   #can't estimate plot intercept effect, variance = 0

# simp2a <- lm.rrpp(Simp ~  Site_short + FungicideTrt +
#                       Site_short:FungicideTrt + Site_short:FungicideTrt:PlotID,
#                  data = samps, iter = 999, print.progress = FALSE, SS.type="III")   
# #anova(rrpp1a, effect.type = "F")
# anova(simp2a, effect.type = "F", 
#      error = c("Site_short:FungicideTrt:PlotID", "Site_short:FungicideTrt:PlotID",
#                "Site_short:FungicideTrt:PlotID", "Residuals"))
#                                Df      SS      MS     Rsq      F        Z Pr(>F)  
# Site_short                      1  1.7341 1.73410 0.06046 1.3509  1.00123  0.156  
# FungicideTrt                    1  2.9708 2.97078 0.10357 2.3143  1.32487  0.094 .
# Site_short:FungicideTrt         1  0.0582 0.05817 0.00203 0.0453 -0.93314  0.817  
# Site_short:FungicideTrt:PlotID  8 10.2693 1.28367 0.35804 2.2517  1.66847  0.042 *
# Residuals                      24 13.6824 0.57010 0.47703                         
# Total                          35 28.6825                                         


range(samps$Simp) #inverse Simpsons
tapply(samps$Simp, samps$Site, mean)
tapply(samps$Simp, samps$FungicideTrt, mean)
```



### Figures
```{r, echo = FALSE, results = 'hide'}
samps %>% select(Sample_ID, IF, Sdiv, Hdiv, Simp, Trt, Site) %>% 
    pivot_longer(cols = c("IF", "Sdiv", "Hdiv", "Simp"), 
                 names_to = "Measure", values_to = "Value") -> samps_w
samps_w$Measure <- factor(samps_w$Measure, levels = c("IF", "Sdiv", "Hdiv", "Simp"))

my_labeller <- c('IF' = "a)  Isolation Frequency",
                 'Sdiv' = "b)  Observed Richness",
                 'Hdiv' = "c)  Shannon Diversity",
                 'Simp' = "d)  Inverse Simpson")

fungPanel <- ggplot(samps_w, aes(x = Trt, y = Value, fill = Site)) +
    facet_grid(Measure~., scales = "free_y", 
               labeller = labeller(Measure = my_labeller)) +
    geom_boxplot(outlier.shape = NA) + 
    #geom_violin() +
    geom_jitter(size = 2, width =.15) +
    scale_y_continuous("") +
    scale_x_discrete("Site-Treatment") +
    scale_fill_manual("", values = c("#66C2A5", "#FC8D62")) +
    theme(axis.text.x  = element_text(angle=60, vjust=0.5, size=10),
          legend.position="none", axis.title.y = element_blank())
#tiff("./Fig3_IF_Richness_trt.tiff", width=3.25, height=9, units="in", res=600)
fungPanel
#dev.off()
```

```{r, include=FALSE}
# boxplots, individual
fungIF <- ggplot(samps, aes(x = Trt, y = IF, fill = Site)) +
    geom_boxplot(outlier.shape = NA) + geom_jitter(size = 2, width =.1) +
    scale_y_continuous("Fungal Isolation Frequency") +
    scale_x_discrete("Site-Treatment") +
    scale_fill_manual("", values = c("#66C2A5", "#FC8D62")) +
    theme(axis.text.x  = element_text(angle=60, vjust=0.5, size=10),
          legend.position="none")
#tiff("./IF_trt.tiff", width=3.25, height=4, units="in", res=600)
fungIF
#dev.off()
rich <- ggplot(samps, aes(x = Trt, y = Sdiv, fill = Site)) +
    geom_boxplot(outlier.shape = NA) + geom_jitter(size = 2, width =.1) +
    scale_y_continuous("Fungal Richness") +
    scale_x_discrete("Site-Treatment") +
    scale_fill_manual("", values = c("#66C2A5", "#FC8D62")) +
    theme(axis.text.x  = element_text(angle=60, vjust=0.5, size=10),
          legend.position="none")
#tiff("./Richness_trt.tiff", width=3.25, height=4, units="in", res=600)
rich
#dev.off()

# barplots, individual
# ggplot(samp_means, aes(x = Trt, y = IF_mean, fill = Site)) +
#     geom_bar(stat = "identity", color = 'black') +
#     geom_errorbar(aes(ymin=IF_mean-IF_sd, ymax=IF_mean+IF_sd), width=.2) +
#     scale_y_continuous("Fungal Isolation Frequency") +
#     scale_x_discrete("Location & Fungicide Treatment") +
#     scale_fill_manual("", values = c("#66C2A5", "#FC8D62")) +
#     theme(axis.text.x  = element_text(angle=60, vjust=0.5, size=14),
#           legend.position="none")
# ggplot(samp_means, aes(x = Trt, y = Sdiv_mean, fill = Site)) +
#     geom_bar(stat = "identity", color = 'black') +
#     geom_errorbar(aes(ymin=Sdiv_mean-Sdiv_sd, ymax=Sdiv_mean+Sdiv_sd), width=.2) +
#     scale_y_continuous("Fungal Richness") +
#     scale_x_discrete("Location & Fungicide Treatment") +
#     scale_fill_manual("", values = c("#66C2A5", "#FC8D62")) +
#     theme(axis.text.x  = element_text(angle=60, vjust=0.5, size=14),
#           legend.position="none")

```

### Figures - Singletons barplot
```{r}
single_bar <- ggplot(samps, aes(x = Trt, y = Singles, fill = FungicideTrt)) +
    geom_col() +
    labs(y = "Count of Singletons", x = "Site-Treatment") +
    #scale_x_discrete("Site-Treatment") +
    scale_fill_manual("", values = c("#8DA0CB","#E78AC3")) +
    theme(axis.text.x  = element_text(angle=60, vjust=0.5, size=10),
          legend.position="none")

#tiff("./SingletonCount.tiff", width=3.25, height=4, units="in", res=600)
single_bar
#dev.off()

```



# Structure
### Models

#### NO nested block random effect
```{r}
rrpp1 <- lm.rrpp(vegdist(SbyS.hel, method = "bray") ~  Site_short*FungicideTrt, 
                 data = samps, iter = 9999, print.progress = FALSE, SS.type="III")   
anova(rrpp1, effect.type = "F")
#                   Df      SS      MS     Rsq      F       Z Pr(>F)  
# Site               1  0.4943 0.49426 0.03612 1.3684 1.10656  0.143  
# FungicideTrt       1  0.3887 0.38869 0.02841 1.0761 0.40707  0.344  
# Site:FungicideTrt  1  0.5779 0.57792 0.04223 1.6000 1.46680  0.069 .
# Residuals         32 11.5584 0.36120 0.84470                        
# Total             35 13.6835             
```

#### YES nested block random effect
```{r}
rrpp1a <- lm.rrpp(vegdist(SbyS.hel, method = "bray") ~  Site_short + FungicideTrt +
                      Site_short:FungicideTrt + Site_short:FungicideTrt:PlotNum,   
                 data = samps, iter = 999, print.progress = FALSE, SS.type="III")   
#anova(rrpp1a, effect.type = "F")
anova(rrpp1a, effect.type = "F", 
     error = c("Site_short:FungicideTrt:PlotNum", "Site_short:FungicideTrt:PlotNum",
               "Site_short:FungicideTrt:PlotNum", "Residuals"))
#                                 Df      SS      MS     Rsq      F        Z Pr(>F)  
# Site_short                       1  0.6166 0.61663 0.04506 1.3798  1.90000  0.032 *
# FungicideTrt                     1  0.3296 0.32957 0.02409 0.7375  0.12715  0.451  
# Site_short:FungicideTrt          1  0.2418 0.24182 0.01767 0.5411 -0.74490  0.756  
# Site_short:FungicideTrt:PlotNum  8  3.5752 0.44690 0.26128 1.3435  2.03221  0.023 *
# Residuals                       24  7.9832 0.33263 0.58342                         
# Total                           35 13.6835                                         
```


#### Models split by location
```{r}
samps_C <- samps[samps$Site_short=="C",] %>% droplevels
samps_G <- samps[samps$Site_short=="G",] %>% droplevels
SbyS.hel_C <- SbyS.hel[rownames(SbyS.hel) %in% samps_C$Sample_ID,]
SbyS.hel_C <- SbyS.hel_C[,colSums(SbyS.hel_C)>0] #drop OTU13 and 14
SbyS.hel_G <- SbyS.hel[rownames(SbyS.hel) %in% samps_G$Sample_ID,]
SbyS.hel_G <- SbyS.hel_G[,colSums(SbyS.hel_G)>0] #drop OTU12


### NO nested block random effect
rrpp1_C <- lm.rrpp(vegdist(SbyS.hel_C, method = "bray") ~  FungicideTrt, 
                 data = samps_C, iter = 9999, print.progress = FALSE, SS.type="III")   
anova(rrpp1_C, effect.type = "F")
#              Df    SS      MS    Rsq      F       Z Pr(>F)
# FungicideTrt  1 0.263 0.26298 0.0687 1.1803 0.63438  0.254
# Residuals    16 3.565 0.22281 0.9313                      
# Total        17 3.828                                     
rrpp1_G <- lm.rrpp(vegdist(SbyS.hel_G, method = "bray") ~  FungicideTrt, 
                 data = samps_G, iter = 9999, print.progress = FALSE, SS.type="III")   
anova(rrpp1_G, effect.type = "F")
#              Df     SS      MS     Rsq      F      Z Pr(>F)  
# FungicideTrt  1 0.5123 0.51227 0.11591 2.0978 1.8073  0.037 *
# Residuals    16 3.9071 0.24419 0.88409                       
# Total        17 4.4194                                       



### YES nested block random effect
rrpp1a_C <- lm.rrpp(vegdist(SbyS.hel_C, method = "bray") ~  FungicideTrt +
                        FungicideTrt:PlotNum,
                 data = samps_C, iter = 999, print.progress = FALSE, SS.type="III")   
#anova(rrpp1a_C, effect.type = "F")
anova(rrpp1a_C, effect.type = "F", 
     error = c("FungicideTrt:PlotNum", "Residuals"))
#                      Df     SS      MS     Rsq      F       Z Pr(>F)  
# FungicideTrt          1 0.2146 0.21456 0.05605 0.7069 0.06591  0.470  
# FungicideTrt:PlotNum  4 1.2140 0.30351 0.31714 1.5492 1.74713  0.041 *
# Residuals            12 2.3510 0.19592 0.61415                        
# Total                17 3.8280                                        
                                        
rrpp1a_G <- lm.rrpp(vegdist(SbyS.hel_G, method = "bray") ~  FungicideTrt +
                        FungicideTrt:PlotNum,
                 data = samps_G, iter = 999, print.progress = FALSE, SS.type="III")   
#anova(rrpp1a_G, effect.type = "F")
anova(rrpp1a_G, effect.type = "F", 
     error = c("FungicideTrt:PlotNum", "Residuals"))
#                      Df     SS      MS     Rsq      F       Z Pr(>F)
# FungicideTrt          1 0.2940 0.29396 0.06652 0.9683 0.64802  0.260
# FungicideTrt:PlotNum  4 1.2144 0.30360 0.27479 1.3530 1.19827  0.118
# Residuals            12 2.6927 0.22439 0.60929                      
# Total                17 4.4194                                      
```


```{r, include = FALSE}
# adonis2
comm1 <- adonis2(SbyS.hel ~  Site_short*FungicideTrt,
                 data = samps, perm=10000, method='bray', by = 'margin' )
comm1
#                   Df SumOfSqs      R2      F   Pr(>F)
# Site               1   0.3704 0.11034 4.4731 0.001999 **
# FungicideTrt       1   0.1629 0.04853 1.9673 0.103632
# Site:FungicideTrt  1   0.1738 0.05176 2.0984 0.076974 .
# Residual          32   2.6501 0.78937
# Total             35   3.3572 1.00000
# comm1_C <- adonis2(SbyS.hel_C ~  FungicideTrt, 
#                  data = samps_C, perm=9999, method='bray' )   
# comm1_C
# #          Df SumOfSqs      R2      F Pr(>F)
# # Model     1  0.09467 0.07657 1.3268 0.2616
# # Residual 16  1.14164 0.92343              
# # Total    17  1.23631 1.00000              
# comm1_G <- adonis2(SbyS.hel_G ~  FungicideTrt, 
#                  data = samps_G, perm=9999, method='bray' )   
# comm1_G
# #          Df SumOfSqs      R2      F  Pr(>F)  
# # Model     1  0.24202 0.13826 2.5671 0.04265 *
# # Residual 16  1.50844 0.86174                 
# # Total    17  1.75047 1.00000                 

```

### Figure
```{r, echo = FALSE, results = 'hide'}
# PCoA
pcoa <- cmdscale(vegdist(SbyS.hel, method = "bray"), eig =TRUE)
explainvar1 <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar1
explainvar2 <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100 
explainvar2

PCoAscores <- as.data.frame(pcoa$points)
PCoAscores$Sample_ID <- rownames(PCoAscores) 
#merge the temporary and SbyE dataframes
samps <- merge(PCoAscores, samps, by="Sample_ID")
rownames(samps) <- samps$Sample_ID

#data for ellipses
ellipses.int <- data.frame() 
for(g in levels(samps$Trt)){
  ellipses.int <- rbind(ellipses.int, cbind(as.data.frame(with(
    samps[samps$Trt==g,],
    veganCovEllipse(cov.wt(cbind(V1,V2),
         wt=rep(1/length(V1),length(V1)))$cov, 
         center=c(mean(V1),mean(V2))))) , Trt=g))   }


ord <- ggplot() +
    coord_equal() +
    geom_path(data = ellipses.int, aes(x = V1, y = V2, 
          group = Trt, colour = Trt, linetype = Trt), linewidth = 1.4,  
          show.legend = TRUE) +
    geom_point(data = samps, aes(x = V1, y = V2, 
          color = Trt, shape = Trt) , size = 4) +
    scale_colour_manual("", values = trt_color) +
    scale_linetype_manual("", values = c(1,2,1,2)) +
    scale_shape_manual("", values = c(15,16,15,16)) +
    scale_x_continuous(paste("PCoA 1 (", explainvar1, "%)", sep = "")) +
    scale_y_continuous(paste("PCoA 2 (", explainvar2, "%)", sep = "")) +
    guides(colour = guide_legend(nrow = 2)) +
    theme(legend.position = "bottom",
          legend.text = element_text(size = 12))

#tiff("./Fig4_Ordination_trt.tiff", width=7, height=7, units="in", res=600)
ord
#dev.off()


# ggplot() +
#     # geom_path(data = ellipses.int, aes(x = V1, y = V2, 
#     #       group = Trt, colour = Trt, linetype = Trt), linewidth = 1.4,  
#     #       show.legend = TRUE) +
#     geom_point(data = samps, aes(x = V1, y = V2, 
#           color = Trt, shape = PlotID) , size = 4) +
#     scale_colour_manual("", values = trt_color) +
#     #scale_linetype_manual("", values = c(1,2,1,2)) +
#     scale_shape_manual("", values = rep(c(15,19,17,18,10,12), 2)) +
#     scale_x_continuous(paste("PCoA 1 (", explainvar1, "%)", sep = "")) +
#     scale_y_continuous(paste("PCoA 2 (", explainvar2, "%)", sep = "")) +
#     theme(legend.position = "bottom")


```



# Indicator Species Analysis
```{r}
myIndic1 <- multipatt(SbyS.noSing, samps$Site, control = how(nperm = 999), func = "IndVal.g")
summary(myIndic1, indvalcomp = TRUE, alpha = 0.2)
# OTU2:Epicoccum and OTU5:Pleosporales are specific to Champaign

myIndic2 <- multipatt(SbyS.noSing, samps$FungicideTrt, control = how(nperm = 999), func = "IndVal.g")
summary(myIndic2, indvalcomp = TRUE, alpha = 0.2)
#  Group Delaro Complete 12oz  #sps.  1 
#                      A      B  stat p.value
# OTU6:Alternaria 0.8333 0.2778 0.481   0.167
```



```{r, include = FALSE}
### Permuted over Plot
# 12 plots total, but 6 per site, and 3 per site*fung trt combo (plot nested in site-fung combo)
# samps$Trt_short <- samps$Trt; 
# levels(samps$Trt_short)[1] <- "CControl"; levels(samps$Trt_short)[2] <- "CFungicide";
# levels(samps$Trt_short)[3] <- "GControl"; levels(samps$Trt_short)[4] <- "GFungicide"

# # fiddling
# getStrata(myPERM)
# getBlocks(myPERM)
# getWithin(myPERM)
# #check(samps, CTRL) # == factorial(36)
# shuffle(samps, control = myPERM)
# 
# # 1.2.3, 10.11.12, 16.17.18 are Fung, while 4.5.6., 7.8.9, and 13.14.15 are Cont
# samps[,c("PlotNum", "Site_short", "Trt_short")]
# samps2 <- samps %>% filter(Site_short == "C") %>% droplevels()

# # permutation analysis
# CTRL <- how(nperm = 999)
# myPERM <- how(nperm = 999,
#               within = Within(type = "free"),
#               plots = Plots(strata = samps$PlotID, type = "free"),
#               blocks = samps$Trt_short,
#               observed = TRUE)


# myIndic1a <- multipatt(SbyS.noSing, samps$Site, 
#         control = myPERM, func = "IndVal.g")
# summary(myIndic1a, indvalcomp = TRUE, alpha = 0.2)
# 
# 
# myIndic1b <- multipatt(SbyS.noSing, samps$Site, 
#         control = CTRL, func = "IndVal.g")
# summary(myIndic1b, indvalcomp = TRUE, alpha = 0.2)

```

### Figure
```{r, results = 'hide', echp = FALSE}
IndValSignif <- c("OTU2:Epicoccum", "OTU5:Pleosporales", "OTU6:Alternaria")

dat3 <- dat2[dat2$variable %in% IndValSignif,]
#dim(dat3)

indval <- ggplot(dat3, aes(x = Trt, y = value, fill = Site)) +
    facet_grid(variable~., scales = "free") +
    geom_violin() +
    geom_jitter(width=.15, height = 0) +
    scale_fill_manual("Names", values = trt_color[c(1,3)]) +
    scale_x_discrete("Site-Treatment") + 
    scale_y_continuous("Fungal Relative Abundance per Plant") +
    theme(axis.text.x  = element_text(angle=60, vjust=0.5, size=10)) +
    guides(fill = "none")
#tiff("./Fig5_Indicator_OTUs.tiff", width=3.25, height=6.25, units="in", res=600)
indval
#dev.off()
```

# Choose what to archive
```{r, include = FALSE}
#samps_otu %>% select(Sample_ID, PlotID, Trt, starts_with("OTU")) -> out
#write.csv(out, "./Data/CornDebrisFungicides_Sample_OTUs_2023-06-07.csv")


# for the most abundant OTUs, will archive 5 replicates per site-Trt combo, if it exists
only5reps <- c("OTU1", "OTU2", "OTU3", "OTU4", "OTU5")

highAbundToArchive <- done %>% filter(OTU %in% only5reps) %>% droplevels() %>% 
    group_by(OTU, Site, FungicideTrt) %>% 
    slice_sample(n=5)
 # really neat funciton from dplyr --> slice_sample to randomly select samples by group
length(highAbundToArchive$Project_Specific_Culture_ID)
#77 isolates (updated now 82)

table(highAbundToArchive$OTU)
table(highAbundToArchive$OTU, highAbundToArchive$FungicideTrt)
table(highAbundToArchive$OTU, highAbundToArchive$Site)


# for the least abundant OTUs and singletons, just archive everything 
lowAbundToArchive <- droplevels(done[done$OTU %nin% only5reps, ])
length(lowAbundToArchive$Project_Specific_Culture_ID)
#48 = 18 singletons (OTU15-OTU32) + 30 isolates to archive from the low abundance OTUs (OTU6-OTU14)
# updated now 51 = 18 + 33


# combine
ToArchive <- rbind(highAbundToArchive, lowAbundToArchive)
ToArchive <- droplevels(ToArchive)
ToArchive <- ungroup(ToArchive)
# 125 = 77 + 48   (updated now 133 = 82 + 51)
ToArchive <- ToArchive[order(ToArchive$Sample_ID, ToArchive$Project_Specific_Culture_ID),]
ToArchive %>% dplyr::select(Project_Specific_Culture_ID, OTU) -> out

#write.csv(out, "./Data/CornDebrisFungicides_ToArchive_2023-06-27.csv")

```




#### end
```{r}
sessionInfo()
```
